---
title: "ESM 262 Assignment 2 - Almond Yield"
author: "Genevieve Chiong, Peter Menzies, Cullen Molitor"
date: "4/9/2022"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE)

library(tidyverse)
library(here)
library(patchwork)
library(kableExtra)

source(here("R/almond_yield.R"))
```

```{r}
# read in the clim data 
clim <- read.table(here("data", "clim_original.txt"))
```

```{r}
# run function 
yield_anomalies <- almond_yield(clim) 

# create table for the clim data 
yield_anomalies %>% 
  kable(col.names = c("Year", "Almond Yield"),
        booktabs = T) %>%
  kable_styling(latex_options = c("striped", "hold_position"), 
                position = "center",
                full_width = FALSE)
```

**Table 1.** Annual almond yield anomaly (tons/acre) estimated by linear regression using precipitation and temperature datafor 1998-2010.

```{r}
# graph yield anomalies for clim dataset
yield_plot <- ggplot(yield_anomalies, aes(x = year, y = yield)) +
  geom_line() +
  labs(title = "Annual Almond Yield Anomalies", 
       y = "Yield (tons/acre)", x = "Year", 
       tag = "a") +
  theme_minimal() +
  theme(plot.title = element_text(hjust = 0.5))
```

```{r}
# find sum of precipitation for January 
sum_precip <- clim %>% 
  filter(month == 1) %>% 
  group_by(year, month) %>% 
  summarize(sum_precip = sum(precip))

# find mean of minimum temperature for February
mean_min_temp <- clim %>% 
  filter(month == 2) %>% 
  group_by(year, month) %>% 
  summarize(mean_min_temp = mean(tmin_c))
```

```{r}
# graph precipitation 
precip_plot <- ggplot(sum_precip, aes(x = year, y = sum_precip)) +
  geom_line() +
  labs(title = "January Precipitation", 
       y = "Precipitation (mm)", x = "Year", 
       tag = "b") +
  theme_minimal() +
  theme(plot.title = element_text(hjust = 0.5))

# graph temperature 
temp_plot <- ggplot(mean_min_temp, aes(x = year, y = mean_min_temp)) +
  geom_line() +
  labs(title = "February Mininum Temperatures", 
       y = "Mean Minimum \n February \n Temperature (ºC)", x = "Year", 
       tag = "c") +
  theme_minimal() +
  theme(plot.title = element_text(hjust = 0.5))
```

```{r}
yield_plot / precip_plot / temp_plot
```


**Graph 1.** Plot (a) shows estimated annual almond yield anomaly based on temperature and precipitation data. 
Plot (b) shows summed precipitation in January of each year. 
Plot (c) shows mean minimum daily temperatures in February of each year.


Our results show the almond yield anomaly from 1988 to 2010 is largely positive. The negative values are very close to 0, while the positive values can be rather large in comparison. There was a spike in almond yield in 1995 which matches up with the spike in precipitation the same year. The precipitation is likely due to the moderate El Niño event in 1994-1995. Temperature and precipitation values were not available for the months in 1988 and therefore there is no yield anomaly available. 


